Sample-Based Determination of Angular Observables

### 2014-06-24 draft.tex
Work on section III.A

\danny{
So far, we assumed that our underlying physical process was accurately described by our PDF $P$.
However, in several interesting processes we might only have an approximative result for $P$.
We will now show that this is the case for the interesting class four-body decays $B\to P_1 P_2 \ell_1 \bar\ell_2$,
which includes the rare $b\to s$-mediated $B$ decay $B \to K\pi\ell^+\ell^-$ and the $V_{ub}$ suppressed decay
$B\to \pi\pi\ell^+\bar\nu_\ell$. For the sake of clarity and simplicity, we will here discuss
the application of our method to the decay $B\to \pi\pi\ell^+\bar\nu_\ell$.\\

We now face the problem that $P$ has a given dependence on the dilepton helicity angle $\theta_1$
and the azimuthal angle $\theta_3$. However, the dipion system can have an arbitrary large
total angular momentum; only the its third component is restricted to $J^{\pi\pi}_z = -1,0,+1$.
Therefore, we will restrict integration to the angles $\theta_1$ and $\theta_3$. \dots
}

@@

The class of $P\to P_1 P_2 \ell_1\bar\ell_2$ decays, with $P,P_1,P_2$ pseudoscalar mesons,
and $\ell_1$, $\bar\ell_2$ charge or neutral leptons, includes decays
such as e.g. $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ and $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$.
Their PDFs can generally be written as \cite{Lee:1992ih}
\begin{equation}
    P(q^2, k^2, \cos\theta_1, \cos \theta_2, \phi) = \frac{1}{\Gamma}\frac{\dd^5 \Gamma}{\dd q^2\,\dd k^2\, \dd\cos\theta_1\,\dd\cos\theta_2\,\dd\phi} = \frac{1}{4\pi} \sum_i \frac{I_i(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\cos\theta_1,\phi)\,,
\end{equation}
where
\begin{equation}
\begin{aligned}
    f_1 & = 1\,,              &
    f_2 & = \cos 2\theta_1\,, &
    f_6 & = \cos  \theta_1\,, \\
    f_3 & = \sin^2\theta_1 \cos 2\phi\,,&
    f_4 & = \sin 2\theta_1 \cos  \phi\,,&
    f_5 & = \sin  \theta_1 \cos  \phi\,,\\
    f_9 & = \sin^2\theta_1 \sin 2\phi\,.&
    f_8 & = \sin 2\theta_1 \sin  \phi\,,&
    f_7 & = \sin  \theta_1 \sin  \phi\,,
\end{aligned}
\end{equation}
Here we denote dilepton mass squared and the dilepton helicity angle as $q^2$ and $\theta_1$, while
the dimeson mass squared and the dimeson helicity angles are denoted as $k^2$ and $\theta_2$. The
azimuthal angle is denoted as $\phi$. The total decay width is
\begin{equation}
    \langle \Gamma\rangle = \int_0^\pi \dd\theta_1\,\frac{\langle 3I_1(\cos\theta_1) - I_2(\cos\theta_1)}{3}\,.
\end{equation}
As usual, we transform from a $\cos\theta_1$-dependence to a $\theta_1$-dependence.
\begin{equation}
\begin{aligned}
    P(q^2, k^2, \theta_1, \theta_2, \phi)
    & = \frac{1}{4\pi} \sum_i \frac{I(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\theta_1, \phi) \sin\theta_1 \sin\theta_2\\
    & \equiv \frac{1}{4\pi} \sum_i P_i(q^2, k^2, \theta_2) f_i(\theta_1, \phi) \sin\theta_2\,.
\end{aligned}
\end{equation}
where we also define $P_i(q^2, k^2, \theta_2) \equiv \theta_2 I_i(q^2, k^2, \cos\theta_2) / \Gamma$.\\

We provide a basis $\hat{f}_i$ of angular functions which is dual to the basis $f_j$ in the sense that
\begin{equation}
    \frac{1}{4\pi}\int_0^\pi \dd\theta_1 \int_0^{2\pi}\dd \phi \hat{f}_i(\theta_1,\phi) f_j(\theta_1,\phi) \sin\theta_1 = \delta_{ij}\,,
\end{equation}
thereby enabling us to project out exactly one of the angular
observables $P_i(q^2)$
\begin{equation}
    \int_0^\pi \dd\theta_1 \int_0^{2\pi}\dd \phi \hat{f}_i(\theta_1,\phi) P(q^2, k^2, \theta_1,\theta_2,\phi) = P_i(q^2, k^2, \theta_2)\,.
\end{equation}
The dual basis reads
\begin{equation}
\begin{aligned}
    \hat{f}_1 & = \frac{21 + 15 \cos 2\theta_1}{16}\,, &
    \hat{f}_2 & = \frac{15 + 45 \cos 2\theta_1}{16}\,, &
    \hat{f}_6 & = 3 \cos\theta_1                   \,, \\
    \hat{f}_3 & = \frac{15}{4}                \cos 2\phi     \,, &
    \hat{f}_4 & = \frac{15}{4} \sin 2\theta_1 \cos  \phi     \,, &
    \hat{f}_5 & = 3            \sin  \theta_1 \cos  \phi     \,, \\
    \hat{f}_9 & = \frac{15}{4}                \sin 2\phi     \,, &
    \hat{f}_8 & = \frac{15}{4} \sin 2\theta_1 \sin  \phi     \,, &
    \hat{f}_7 & = 3            \sin  \theta_1 \sin  \phi     \,.
\end{aligned}
\end{equation}

>>>>

So far, we assumed that our underlying physical process was accurately described by our PDF $P$.
However, in several interesting processes we might only have an approximative result for $P$.
We will now show that this is the case for the interesting class of four-body decays $B\to P_1 P_2 \ell_1 \bar\ell_2$,
which includes the rare $b\to s$-mediated $B$ decay $B \to K\pi\ell^+\ell^-$ and the $V_{ub}$ suppressed decay
$B\to \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in the small-width approximation
and when assuming a pure $P$-wave, resonant final state. An extension to $S-P$ interference has been
studied for $B\to K\pi\ell^+\ell^-$ \cite{Blake:2012xx,Becirevic:2011xx} and for $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$
\cite{Faller:2013xx}.

For this class of decays, we face the problem that $P$ has a given dependence on the dilepton helicity angle $\theta_1$
and the azimuthal angle $\theta_3$. (See also \refapp{btokpill} for details on the angular distribution.)
However, on the level of decay amplitudes the dimeson system can have an arbitrary
large total angular momentum; only the third component of its angular momentum is restricted to $J^{\pi\pi}_z = -1,0,+1$.\\

We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(m)}$
with respect to the remaining angle $\theta_2$. Using the observables $P_k(\vec{\nu}=(q^2, k^2))$ as defined in \refapp{}
the expansion reads
\begin{equation}
    P_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,m}} P_{k,j}(\vec{\nu}) p_{j}^{(m)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where $m=0$ for $n=1,2,6$ and $m=1$ for $n=3,4,5,7,8,9$. The angular observables defined as coefficients of this expansion
have the merit of a well-defined total angular momentum, and thus can be disentangled (see \refapp{partial-waves} for
a derivation). As a consequence of the orthogonality of the Legendre polynomials, mismodelling or lack of modelling of higher
partial-wave observables \emph{does not affect} the method of moments as given in the previous section.
Unfortunately, this benefit on the experimental side is accompanied with a theoretical draw back. Each observable
$P_{k,j}(\vec{\nu})$ consists of an infinite sum of bilinear of partial-wave amplitudes. It remains for theoretical
analyses to estimate or calculate the impact of partial waves beyond the P wave \cite{Das:2014}.\\

We wish to emphasize, however, that detector acceptance effects systematically affect the expansion for any basis
of angular functions. The means to cope with this problem are discussed in the following subsection.

@@

The class of $P\to P_1 P_2 \ell_1\bar\ell_2$ decays, with $P,P_1,P_2$ pseudoscalar mesons,
and $\ell_1$, $\bar\ell_2$ charge or neutral leptons, includes decays
such as e.g. $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ and $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$.
Their PDFs can generally be written as \cite{Lee:1992ih}
\begin{equation}
    P(q^2, k^2, \cos\theta_1, \cos \theta_2, \phi) = \frac{1}{\langle \Gamma\rangle}\frac{\dd^5 \Gamma}{\dd q^2\,\dd k^2\, \dd\cos\theta_1\,\dd\cos\theta_2\,\dd\phi} = \frac{1}{4\pi} \sum_i \frac{I_i(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\cos\theta_1,\phi)\,,
\end{equation}
where
\begin{equation}
\begin{aligned}
    f_1 & = 1\,,              &
    f_2 & = \cos 2\theta_1\,, &
    f_6 & = \cos  \theta_1\,, \\
    f_3 & = \sin^2\theta_1 \cos 2\phi\,,&
    f_4 & = \sin 2\theta_1 \cos  \phi\,,&
    f_5 & = \sin  \theta_1 \cos  \phi\,,\\
    f_9 & = \sin^2\theta_1 \sin 2\phi\,.&
    f_8 & = \sin 2\theta_1 \sin  \phi\,,&
    f_7 & = \sin  \theta_1 \sin  \phi\,,
\end{aligned}
\end{equation}
Here we denote dilepton mass squared and the dilepton helicity angle as $q^2$ and $\theta_1$, while
the dimeson mass squared and the dimeson helicity angles are denoted as $k^2$ and $\theta_2$. The
azimuthal angle is denoted as $\phi$. The total decay width is
\begin{equation}
    \langle \Gamma\rangle = \int_0^\pi \dd\theta_1\,\dd q^2\,\dd k^2\,\sin\theta_2 \frac{3I_1(q^2, k^2, \cos\theta_2) - I_2(q^2, k^2, \cos\theta_2)}{3}\,.
\end{equation}
As usual, we transform from a $\cos\theta_1$-dependence to a $\theta_1$-dependence.
\begin{equation}
\begin{aligned}
    P(q^2, k^2, \theta_1, \theta_2, \phi)
    & = \frac{1}{4\pi} \sum_i \frac{I(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\theta_1, \phi) \sin\theta_1 \sin\theta_2\\
    & \equiv \frac{1}{4\pi} \sum_i P_i(q^2, k^2, \theta_2) f_i(\theta_1, \phi)\,.
\end{aligned}
\end{equation}
where we also define $P_i(q^2, k^2, \theta_2) \equiv \sin\theta_2 I_i(q^2, k^2, \cos\theta_2) / \Gamma$.\\

We provide a basis $\hat{f}_i$ of angular functions which is dual to the basis $f_j$ in the sense that
\begin{equation}
    \frac{1}{4\pi}\int_0^\pi \dd\theta_1 \int_0^{2\pi}\dd \phi \hat{f}_i(\theta_1,\phi) f_j(\theta_1,\phi) \sin\theta_1 = \delta_{ij}\,,
\end{equation}
thereby enabling us to project out exactly one of the angular
observables $P_i(q^2, k^2, \theta_2)$
\begin{equation}
    \int_0^\pi \dd\theta_1 \int_0^{2\pi}\dd \phi \hat{f}_i(\theta_1,\phi) P(q^2, k^2, \theta_1,\theta_2,\phi) = P_i(q^2, k^2, \theta_2)\,.
\end{equation}
The dual basis reads
\begin{equation}
\begin{aligned}
    \hat{f}_1 & = \frac{21 + 15 \cos 2\theta_1}{16}\,, &
    \hat{f}_2 & = \frac{15 + 45 \cos 2\theta_1}{16}\,, &
    \hat{f}_6 & = 3 \cos\theta_1                   \,, \\
    \hat{f}_3 & = \frac{15}{4}                \cos 2\phi     \,, &
    \hat{f}_4 & = \frac{15}{4} \sin 2\theta_1 \cos  \phi     \,, &
    \hat{f}_5 & = 3            \sin  \theta_1 \cos  \phi     \,, \\
    \hat{f}_9 & = \frac{15}{4}                \sin 2\phi     \,, &
    \hat{f}_8 & = \frac{15}{4} \sin 2\theta_1 \sin  \phi     \,, &
    \hat{f}_7 & = 3            \sin  \theta_1 \sin  \phi     \,.
\end{aligned}
\end{equation}
